The galactic chemical evolution of carbon: Implications for stellar nucleosynthesis

### 2025-11-19 carbon_paper/tex/main.tex
james grammar edits

\newcommand{\fruity}{\texttt{\hyperlink{fruity}{Fruity}}}
\newcommand{\nugrid}{\texttt{\hyperlink{nugrid}{NuGrid}}}
\newcommand{\monash}{\texttt{\hyperlink{monash}{Monash}}}
\newcommand{\aton}{\texttt{\hyperlink{aton}{ATON}}}
\newcommand{\cfactor}{2.57}
\newcommand{\zetaobest}{1.90 \pm 0.10}
\newcommand{\zetaibest}{2.28 \pm 0.16}

\newcommand{\nsubgiants}{14,066}

% displays fractions within fractions well
\newcommand{\ddfrac}[2]{\ensuremath{%
    \frac{\displaystyle{#1}}{\displaystyle{#2}}
}}

% Acronyms
\newcommand{\agb}{AGB}
\newcommand{\apogee}{APOGEE}
\newcommand{\aspcap}{\textsc{aspcap}}
\newcommand{\cc}{CCSN}
\newcommand{\gce}{GCE}
\newcommand{\ia}{SN Ia}

% internal abbreviations
\newcommand{\caah}{[C/Mg]-[Mg/H]}
\newcommand{\caafe}{[C/Mg]-[Mg/Fe]}
\newcommand{\oh}{{\rm [O/H]}}
\newcommand{\mgh}{{\rm [Mg/H]}}
\newcommand{\cmg}{{\rm [C/Mg]}}
\newcommand{\co}{{\rm [C/O]}}
\newcommand{\mgfe}{{\rm [Mg/Fe]}}

\newcommand{\Var}{\operatorname{Var}}

\newcommand{\Hii}{H~{\sc ii}}

% other

\makeatletter
\newcommand{\C}[1][\@nil]{
    \def\tmp{#1}%
    \ifx\tmp\@nnil%
        \ensuremath{\rm C}%
    \else%
        \ifmmode ^{#1}{\rm C}%
        \else $^{#1}$C%
        \fi%
\fi }
\makeatother

\newcommand{\Yct}{{y_{\rm C}^{\rm  tot}}}
\newcommand{\Ycc}{{y_{\rm C}^{\rm CC}}}
\newcommand{\Ymg}{{y_{\rm Mg}^{\rm CC}}}
\newcommand{\Ycagb}{{y_{\rm C}^{\rm AGB}}}
\newcommand{\ycagb}{\y_{\rm C}^{\rm AGB}}
\newcommand{\y}{p}

@@

\newcommand{\about}[1]{${\sim} #1$}

@@

% Caption macros
\newcommand{\captionline}[2][very
thick]{\tikz[baseline={([yshift=-.5ex]current bounding box.base)}]{
\draw[#2,#1, line cap=round] (0,0) -- (0.3,0); }}

%%%%%%%%%%% JWJ' editing macros %%%%%%%%%%%
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\strike}[1]{{\color{red} \sout{#1}}}
% \newcommand{\add}[1]{{\color{ForestGreen} #1}}
% \newcommand{\note}[1]{{\color{ForestGreen} \textit{ \small (JWJ: #1)}}}
% \LetLtxMacro\origcitep\citep
% \LetLtxMacro\origcitet\citet
% \renewcommand{\citep}[2][]{%
%     \mbox{\origcitep[#1][]{#2}}
% }
% \renewcommand{\citet}[2][]{%
%     \mbox{\origcitet[#1][]{#2}}
% }

\newcommand{\dbstrike}[1]{{\color{Thistle} \sout{#1} }}
\newcommand{\dbadd}[1]{{\color{Thistle} #1}}
\newcommand{\dbnote}[1]{ {\color{Thistle} \textit{\small (DB: #1)}} }

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{%
Daniel A. Boyea,$^{1, 2, 3}$\thanks{%
Contact e-mail:~\href{mailto:danielaboyea@gmail.com}{danielaboyea@gmail.com}}
James W. Johnson,$^{1, 2, 4}$
and others
%David H. Weinberg,$^{1,2}$
%Jennifer A. Johnson,$^{1,2}$
%Falk Herwig$^{3}$
\\
% List of institutions
$^{1}$Department of Astronomy, The Ohio State University, 140 W. 18th Ave., Columbus, OH, 43210, USA
\\
$^{2}$Center for Cosmology \& Astroparticle Physics (CCAPP), The Ohio State University, 191 W. Woodruff Ave., Columbus, OH, 43210, USA
\\
$^{3}$Department of Physics \& Astronomy, University of Victoria, Victoria, BC V8P 5C2, Canada
\\
$^{4}$The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA, 91101, USA
}

@@

\begin{figure*}
    \centering
    \includegraphics{subgiants.pdf}
    \caption{The [C/Mg] ratio versus [Mg/H] (left) and [Mg/Fe] (right) for the \citet{jack}~sample of \apogee{} subgiants. {\bf Left:} High- and low-$\alpha$ stars are shown in blue and orange, respectively, using the separation defined in Eq.~\ref{eq:high_alpha}. The black points represent the median trend along the low-$\alpha$ sequence. {\bf Right:} Stars are colour-coded by their [Mg/H] abundance. The black points show the median [C/Mg]-[Mg/Fe] sequence for stars in the $-0.15 \leq [{\rm Mg/H}] \leq -0.05$ range. We use these median trends as our empirical benchmark in subsequent figures. } \label{fig:subgiants}
\end{figure*}

@@

Given the challenge of theoretical yield predictions, previous work has attempted to constrain chemical yields empirically. 
Even early chemical evolution work, such as \citet{garnett1997}, began to test stellar theory with observed abundance trends.
More recently, \citet{weinberg+19, weinberg+22} analysed the abundance ratio trends of high- and low-$\alpha$ sequence stars in APOGEE~\citep{apogee17} to infer relative yields of prompt and delayed enrichment sources and their metallicity dependence.
\citet{emily+19, emily+22, emily+24} applied the same methodology to GALAH~\citep{DeSilva2015, Martell2017}. In addition, models such as \citet{eitner+2020} and \citet{dubay+24} have used abundance trends of Fe and related elements to understand the physics and origin of \ia.
Most directly relevant to our analysis, \citet{james+24} applied Milky Way chemical evolution models to N and recommended a yield prescription that reproduces the observed relation between N and O abundances. We  extend their methodology to C in this paper.

@@

In Section~\ref{sec:data_selection} we describe the selection for our subgiant observational sample.
In Section~\ref{sec:nucleosynthesis}, we discuss theoretical estimates of \agb\ and \cc\ C yields.
In Section~\ref{sec:vice}, we describe the details of our GCE models.
In Section~\ref{sec:results}, we present our model predictions and discuss their observationally distinguishable predictions. 
We then infer constraints on yield parameters in Section~\ref{sec:mcmc}.
In Section~\ref{sec:gas}, we compare our model to observations of halo stars and extragalactic \Hii{} regions from the literature. 
We summarize our conclusions in Section~\ref{sec:conclusions}.

@@

We use the sample of \apogee\ DR17 subgiants from \citet{jack}. We select stars from the ASPCAP pipeline (revision 1) within the $\log g$-$T_\text{eff}$ polygon given by
 \begin{equation}
    \begin{cases} \label{eq:subgiant_selection}
        \log \text{g} \geq 3.5 \\
        \log \text{g} \leq 0.004\,T_{\rm eff} - 15.7 \\
        \log \text{g} \leq 0.0007\,T_{\rm eff} + 0.36 \\
        \log \text{g} \leq -0.0015\,T_{\rm eff} + 12.05 \\
        \log \text{g} \geq 0.0012\,T_{\rm eff} - 2.8. \\
    \end{cases}
\end{equation}
Following \citet{jack}, we also exclude stars with the flags:
        \verb|APOGEE_MIRCLUSTER_STAR|,
        \verb|APOGEE_EMISSION_STAR|,
        \verb|APOGEE_EMBEDDEDCLUSTER_STAR|,
        \verb|APOGEE2_YOUNG_CLUSTER|,
        \verb|APOGEE2_W345|,
        \verb|APOGEE2_EB|, and
        \verb|DUPLICATE|.
We additionally remove stars with no reported (or flagged unreliable) C, N, Mg, or Fe abundances. These cuts yield a sample containing \nsubgiants\ subgiants.

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{The net fractional C yield from AGB stars as a function of initial stellar mass and metallicity. Each panel represents yields from one of four \agb\ studies from the literature: \fruity{}, \aton{}, \monash{}, \nugrid{} (see Section \ref{sec:agb}).  The black dotted line shows $\Ycagb=0$ for reference. 
        }

@@

We explore four different AGB star yield tables from the literature, which provide well-sampled grids in mass and metallicity. We refer to the tables as 
\begin{description}
    \item \hypertarget{fruity}{\texttt{FRUITY}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{aton}{\texttt{ATON}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{monash}{\texttt{Monash}}: \citet{KL16, karakas+18}
    \item \hypertarget{nugrid}{\texttt{NuGrid}}: \citet{pignatari+16, ritter+18, battino+19, battino+21}
\end{description}
Each AGB yield table reports yields for stellar models on a grid of masses and metallicities as reported in Table~\ref{tab:agb}.
We use the same tables as \citet{james+23}, except that we have swapped \citet{karakas10} for \citet{pignatari+16}. (See Appendix~\ref{sec:nugrid_yields} for the implementation and recalculation of these yields.)
For our models to better match observations, we uniformly amplify the yield tables according to
\begin{equation} \label{eq:alpha}
        \ycagb(m, Z) \rightarrow \aagb\ \ycagb(m, Z).
\end{equation}
As our fiducial \agb\ yield, we use the \fruity\ table with $\aagb=\cfactor$, (balancing MW and gas-phase constraints; see Section~\ref{sec:results} and Section~\ref{sec:gas}).

@@

The IMF-averaged yields computed from the literature tables differ in normalization and metallicity dependence.  
The left panel of Fig.~\ref{fig:agb-ssp} shows IMF-averaged C yields for each \agb\ model as a function of metallicity, computed at time $\tau = 10$ Gyr after the formation of a single stellar population.
The normalizations span a factor of \about{2} (e.g., between $6\times 10^{-4}$ and $12 \times 10^{-4}$ at $Z=0.33\,\Zo$).
Additionally, the slope varies by a factor of ${\sim}3$.
Both \nugrid{} and \aton{} predict non-monotonic C yields with $Z$, but modelling systematics may exceed the scale of non-monotonic features.
These variations are due to different choices of reaction rates, convection treatments, and mass-loss rates (see discussion below and \citealt{james+23}).

The right panel of Fig.~\ref{fig:agb-ssp} shows the time-dependence of C production by \agb\ stars in a single stellar population of metallicity ${\rm [M/H]} = -0.1$ as a function of age. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is predicted to be most important for C production, half the yield is produced before \about{1}\,Gyr, similar to Fe production by \ia. 
\monash{} weights C production more heavily towards high-mass \agb\ stars, resulting in shorter delay times, whereas the \fruity\ model predicts a slightly longer time-scale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. Fe production, in contrast, continues steadily for 10\,Gyr.

@@

Both HBB and TDU result in mass and metallicity-dependent C yields. 
Lower mass stars ($\lesssim 1\,\Mo$)  do not experience (strong) TDU. As a result, C yields from these stars are affected only by first dredge-up, resulting in small net C yields \citep[table 1]{karakas10}.
Above ${\sim}1\,\Mo$, TDU becomes more important, enriching the outer layers with C.
In \agb\ stars more massive than ${\sim}5\,\Mo$, efficient HBB turns most of the \C[12] into $^{14}$N.
TDU is more efficient for compact, massive cores. HBB requires a more massive AGB star to reach sufficient temperatures at the base of the convective envelope to initiate the CNO cycle. Both processes are more efficient at low metallicity due to the lower opacities, hotter temperatures, and more compact internal structures. The details of each process are ultimately set by uncertain choices in stellar evolution models.

@@

\begin{table*}
	\centering
    \caption[]{For each \agb\ yield set, the IMF-averaged \agb\ C yield at solar metallicity $y_{\rm C, 0}^{\rm AGB}$, the fraction of solar C produced in the model $f_\odot^{\rm AGB}$, and the masses and metallicities each yield table is sampled on.
    $f_{\sun}^{\rm AGB}$ is calculated based on $y_{\rm C, \sun}^{\rm AGB}$ assuming the fiducial total C yield of $\Yct = 26.7\times 10^{-4}$.
    }

	\label{tab:agb}
    \begin{tabular}{c  c  c p{4cm} p{4cm}} % four columns, alignment for each
		\hline 
        \agb\ table 
                & $y_{\rm C, \sun}^{\rm AGB}\times10^4$ %& y
                &  $f_{\sun}^{\rm AGB}$
                & masses ($\Mo$) & metallicities ($Z$)\\
        \hline
        \fruity 
                & 3.11
                & 0.12
                & 1.3, 1.5, 2, 2.5, 3, 4, 5, 6
                & 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02
                \\
        \aton 
                & -0.009
                & -0.0003
                & 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 6, 6.5, 7
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
        \monash 
                &  2.96
                & 0.11& 1, 1.25, 1.5, 1.75, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.5, 5, 5.5, 6, 7 
                & 0.0028, 0.007, 0.014, 0.03
                \\
        \nugrid 
                & 9.25
                &  0.35
                & 1, 1.65, 2, 3, 4, 5, 6, 7
                &  0.0001, 0.001, 0.006, 0.01, 0.02
                \\
		\hline
	\end{tabular}
\end{table*}

@@

\cc{} models predict a wide range of C yields, spanning nearly a factor of ten, even at fixed metallicity.
Fig.~\ref{fig:y_cc} shows the integrated yields of massive star models from the literature. 
 The \cite{LC18} models, which include rotation, show that the induced mixing (e.g., \citealt{frischknecht+16}) can dramatically increase the magnitude and metallicity dependence of $\Ycc$. As we will later empirically show, \cc\ C production needs to be strongly metallicity-dependent at $Z \approx \Zo$, which is consistent with \citepos{LC18} rapidly rotating models and to a lesser extent \citet{NKT13}.
Fig.~\ref{fig:y_cc} also shows the IMF-averaged \agb\ star yields from the \fruity\ model (in Fig.~\ref{fig:agb-ssp}). Especially at $Z\approx \Zo$, most \cc\ models dominate \agb\ C production.

@@

The metallicity dependence of CCSN C yields is due to an interplay between rotation, opacity, and stellar evolution. As opacity increases with metallicity, massive stars experience stronger winds. In particular, C-enriched envelope material is lost through winds before conversion into heavier elements, resulting in metallicity-dependent C production \citep{LC18}.
While there are many stellar models providing predictions of \cc{} yields, the results of these models are highly uncertain. Furthermore, rotation, binarity, and explodability introduce substantial uncertainties in \cc\ predictions \citep{farmer+21}.

@@

SN Ia produce negligible C. Of the models from \citet{iwamoto+99}, \citet{seitenzahl+13}, and \citet{gronow+21a, gronow+21b}, the maximum C production is less than 4 per cent of our required solar C yield, and most likely less than 2 per cent (assuming  $2.2\times 10^{-3}$ SN Ia explosions per solar mass of star formation; \citealt{maoz+mannucci12}).
Novae are important contributors of \C[13], but do not process enough material to impact \C[12] production \citep{starrfield+20}. \C[13] is beyond the scope of this work and accounts for a small fraction of the C in the Universe.

@@

Fig.~\ref{fig:c_evo} shows evolutionary tracks in \caah{} and \caafe{} in the fiducial model at 1\,kpc intervals of radius.
The \caah\ trend reaches an equilibrium state within \about{5}\,Gyr. \caafe\ tracks continue to evolve due to the long tail of the \ia\ DTD.

The evolution of chemical abundances can be understood as the interplay between different processes with different time-scales. 
 Initially, \cc\ dominate enrichment. The [C/Mg] abundance ratio is set by the relative yield $\Ycc/\Ymg$. Due to the metallicity dependence of $\Ycc$, [C/Mg] increases with  [Mg/H]. [Mg/H], [Fe/H], and [C/H] all rise rapidly. [Mg/Fe] is initially high due to the delayed production time-scales of SN Ia Fe, as expected from the analytic models of \citet{WAF17}.
 Of order 100 Myr later, delayed sources enrich the ISM. \agb\ stars release C, raising [C/Mg], and \ia\ expel Fe, lowering [Mg/Fe]. 
 Several Gyr later, [Mg/H] reaches equilibrium. As the metal abundance stabilizes, so do metallicity-dependent \cc{} yields, allowing C to reach equilibrium. [Fe/H] continues to rise slowly due to ongoing \ia\ from old stellar populations. 
 Finally, the system reaches approximate equilibrium and evolution slows. \

@@

\caption[]{The median low-$\alpha$ [C/N] ratio as a function of [Mg/H] for our fiducial model (blue line), compared with \apogee{} subgiants (medians as black points with 16th and 84th percentile ranges as grey bars). Stars are binned into 20 equal-number bins. Combining our results with the suggested N yield from \citet{james+23} \ (rescaled to our adopted yields) explains the thin-disk evolution of both C and N. 
}
\label{fig:nitrogen}
\end{figure}

@@

    \caption[]{
        Trends in [C/Mg] with [Mg/H] (left) or [Mg/Fe] (right). 
        Stars are binned into 20 (left) or 12 (right) equal-number bins. 
        The left panel only shows low-$\alpha$ stars, and the right panel shows stars where $-0.15\leq \mgh \leq -0.05$.
        Coloured lines represent the median [C/Mg] in bins of [Mg/H] or [Mg/Fe] for each model.
        Black points and grey dashes represent the median and 16th-84th percentiles of [C/Mg] in each bin in the \citet{jack}~sample. 
        The total solar C yield is held fixed in each model. 
        %
        \textbf{Top}: Models with different slopes of the metallicity dependence of the CCSN yield, $\zcc$ (Eq.~\ref{eq:y_cc}), where higher $\zcc$ corresponds to a more rapid increase in C production with metallicity. \textbf{Middle}: Models with different fractions of solar C produced in AGB stars, $\fagb$. 
        \textbf{Bottom:} Models where both $\fagb$ and $\zcc$ are adjusted simultaneously. 
    }
    \label{fig:zeta_f}
\end{figure*}

@@

\section{Model parameters and details}

\begin{table}
	\centering
        \caption[]{Description of the yields used for models presented in this paper.}
	\label{tab:model_parameters}

>>>>

% AGB yield tables
\newcommand{\fruity}{\texttt{\hyperlink{fruity}{Fruity}}}
\newcommand{\nugrid}{\texttt{\hyperlink{nugrid}{NuGrid}}}
\newcommand{\monash}{\texttt{\hyperlink{monash}{Monash}}}
\newcommand{\aton}{\texttt{\hyperlink{aton}{ATON}}}
\newcommand{\cfactor}{2.57}
\newcommand{\nsubgiants}{14,066}

% internal abbreviations
\newcommand{\caah}{[C/Mg]-[Mg/H]}
\newcommand{\caafe}{[C/Mg]-[Mg/Fe]}
\newcommand{\co}{{\rm [C/O]}}
\newcommand{\Hii}{H~{\sc ii}}

% mathmode shortcuts
\newcommand{\Yct}{{y_{\rm C}^{\rm  tot}}}
\newcommand{\Ycc}{{y_{\rm C}^{\rm CC}}}
\newcommand{\Ymg}{{y_{\rm Mg}^{\rm CC}}}
\newcommand{\Ycagb}{{y_{\rm C}^{\rm AGB}}}
\newcommand{\ycagb}{\y_{\rm C}^{\rm AGB}}
\newcommand{\y}{p}

@@

% other
\newcommand{\about}[1]{${\sim} #1$}

@@

%%%%%%%%%%% JWJ' editing macros %%%%%%%%%%%
\newcommand{\red}[1]{{\color{red} #1}}
\newcommand{\strike}[1]{{\color{red} \sout{#1}}}

@@

% The list of authors, and the short list which is used in the headers.
% If you need two or more lines of authors, add an extra line using \newauthor
\author[D. A. Boyea et. al.]{%
Daniel A. Boyea,$^{1, 2}$\thanks{%
Contact e-mail:~\href{mailto:danielaboyea@gmail.com}{danielaboyea@gmail.com}}
James W. Johnson,$^{2, 3, 4}$
and
David H. Weinberg$^{2,4}$
\\
% List of institutions
$^{1}$Department of Physics \& Astronomy, University of Victoria, 3800 Finnerty Road, Victoria, BC, V8P 5C2, Canada
\\
$^{2}$Department of Astronomy, The Ohio State University, 140 W. 18th Ave., Columbus, OH, 43210, USA
\\
$^{3}$The Observatories of the Carnegie Institution for Science, 813 Santa Barbara St., Pasadena, CA, 91101, USA
\\
$^{4}$Center for Cosmology \& Astroparticle Physics (CCAPP), The Ohio State University, 191 W. Woodruff Ave., Columbus, OH, 43210, USA
}

@@

\begin{figure*}
    \centering
    \includegraphics{subgiants.pdf}
    \caption{The [C/Mg] ratio versus [Mg/H] (left) and [Mg/Fe] (right) for the \citet{jack}~sample of APOGEE subgiants. {\bf Left:} High- and low-$\alpha$ stars are shown in blue and orange, respectively, using the separation defined in Eq.~\ref{eq:high_alpha}. The black points represent the median trend along the low-$\alpha$ sequence. {\bf Right:} Stars are colour-coded by their [Mg/H] abundance. The black points show the median [C/Mg]-[Mg/Fe] sequence for stars in the $-0.15 \leq [{\rm Mg/H}] \leq -0.05$ range. We use these median trends as our empirical benchmark in subsequent figures. } \label{fig:subgiants}
\end{figure*}

@@

Given the challenge of theoretical yield predictions, previous work has attempted to constrain chemical yields empirically. 
Even early chemical evolution work, such as \citet{garnett1997}, began to test stellar theory with observed abundance trends.
More recently, \citet{weinberg+19, weinberg+22} analysed the abundance ratio trends of high- and low-$\alpha$ sequence stars in APOGEE~\citep{apogee17} to infer relative yields of prompt and delayed enrichment sources and their metallicity dependence.
\citet{emily+19, emily+22, emily+24} applied the same methodology to GALAH~\citep{DeSilva2015, Martell2017}. In addition, models such as \citet{eitner+2020} and \citet{dubay+24} have used abundance trends of Fe and related elements to understand the physics and origin of SN Ia.
Most directly relevant to our analysis, \citet{james+24} applied Milky Way chemical evolution models to N and recommended a yield prescription that reproduces the observed relation between N and O abundances. We  extend their methodology to C in this paper.

@@

In Section~\ref{sec:data_selection} we describe the selection for our subgiant observational sample.
In Section~\ref{sec:nucleosynthesis}, we discuss theoretical estimates of AGB and CCSN C yields.
In Section~\ref{sec:vice}, we describe the details of our GCE models.
In Section~\ref{sec:results}, we present our model predictions and discuss their observationally distinguishable predictions. 
We then infer constraints on yield parameters in Section~\ref{sec:mcmc}.
In Section~\ref{sec:gas}, we compare our model to observations of halo stars and extragalactic \Hii{} regions from the literature. 
We summarize our conclusions in Section~\ref{sec:conclusions}.

@@

We use the sample of APOGEE DR17 subgiants from \citet{jack}. We select stars from the {\sc aspcap} pipeline (revision 1) within the $\log g$-$T_\text{eff}$ polygon given by
 \begin{equation}
    \begin{cases} \label{eq:subgiant_selection}
        \log \text{g} \geq 3.5 \\
        \log \text{g} \leq 0.004\,T_{\rm eff} - 15.7 \\
        \log \text{g} \leq 0.0007\,T_{\rm eff} + 0.36 \\
        \log \text{g} \leq -0.0015\,T_{\rm eff} + 12.05 \\
        \log \text{g} \geq 0.0012\,T_{\rm eff} - 2.8. \\
    \end{cases}
\end{equation}
Following \citet{jack}, we also exclude stars with the flags:
        \verb|APOGEE_MIRCLUSTER_STAR|,
        \verb|APOGEE_EMISSION_STAR|,
        \verb|APOGEE_EMBEDDEDCLUSTER_STAR|,
        \verb|APOGEE2_YOUNG_CLUSTER|,
        \verb|APOGEE2_W345|,
        \verb|APOGEE2_EB|, and
        \verb|DUPLICATE|.
We additionally remove stars with no reported (or flagged unreliable) C, N, Mg, or Fe abundances. These cuts yield a sample containing \nsubgiants\ subgiants.

@@

\begin{figure*}
    \centering
 	    \includegraphics[scale=1]{agb_yields.pdf}
        \caption[]{The net fractional C yield from AGB stars as a function of initial stellar mass and metallicity. Each panel represents yields from one of four AGB studies from the literature: \fruity{}, \aton{}, \monash{}, \nugrid{} (see Section \ref{sec:agb}).  The black dotted line shows $\Ycagb=0$ for reference. 
        }

@@

We explore four different AGB star yield tables from the literature, which provide well-sampled grids in mass and metallicity. We refer to the tables as 
\begin{description}
    \item \hypertarget{fruity}{\texttt{FRUITY}}: \citet{cristallo+11, cristallo+15}
    \item \hypertarget{aton}{\texttt{ATON}}: \citet{ventura+13,ventura+14,ventura+18, ventura+20}
    \item \hypertarget{monash}{\texttt{Monash}}: \citet{KL16, karakas+18}
    \item \hypertarget{nugrid}{\texttt{NuGrid}}: \citet{pignatari+16, ritter+18, battino+19, battino+21}
\end{description}
Each AGB yield table reports yields for stellar models on a grid of masses and metallicities as reported in Table~\ref{tab:agb}.
We use the same tables as \citet{james+23}, except that we have swapped \citet{karakas10} for \citet{pignatari+16}. (See Appendix~\ref{sec:nugrid_yields} for the implementation and recalculation of these yields.)
For our models to better match observations, we uniformly amplify the yield tables according to
\begin{equation} \label{eq:alpha}
        \ycagb(m, Z) \rightarrow \aagb\ \ycagb(m, Z).
\end{equation}
As our fiducial AGB yield, we use the \fruity\ table with $\aagb=\cfactor$, (balancing MW and gas-phase constraints; see Section~\ref{sec:results} and Section~\ref{sec:gas}).

@@

The IMF-averaged yields computed from the literature tables differ in normalization and metallicity dependence.  
The left panel of Fig.~\ref{fig:agb-ssp} shows IMF-averaged C yields for each AGB model as a function of metallicity, computed at time $\tau = 10$ Gyr after the formation of a single stellar population.
The normalizations span a factor of \about{2} (e.g., between $6\times 10^{-4}$ and $12 \times 10^{-4}$ at $Z=0.33\,\Zo$).
Additionally, the slope varies by a factor of ${\sim}3$.
Both \nugrid{} and \aton{} predict non-monotonic C yields with $Z$, but modelling systematics may exceed the scale of non-monotonic features.
These variations are due to different choices of reaction rates, convection treatments, and mass-loss rates (see discussion below and \citealt{james+23}).

The right panel of Fig.~\ref{fig:agb-ssp} shows the time-dependence of C production by AGB stars in a single stellar population of metallicity ${\rm [M/H]} = -0.1$ as a function of age. 
As the mass range $2\,\Mo\lesssim M \lesssim 4\,\Mo$ is predicted to be most important for C production, half the yield is produced before \about{1}\,Gyr, similar to Fe production by SN Ia. 
\monash{} weights C production more heavily towards high-mass AGB stars, resulting in shorter delay times, whereas the \fruity\ model predicts a slightly longer time-scale of \about{1}\,Gyr. In any case, little to no C is produced more than 2\,Gyr after a star formation event. Fe production, in contrast, continues steadily for 10\,Gyr.

@@

Both HBB and TDU result in mass and metallicity-dependent C yields. 
Lower mass stars ($\lesssim 1\,\Mo$)  do not experience (strong) TDU. As a result, C yields from these stars are affected only by first dredge-up, resulting in small net C yields \citep[table 1]{karakas10}.
Above ${\sim}1\,\Mo$, TDU becomes more important, enriching the outer layers with C.
In AGB stars more massive than ${\sim}5\,\Mo$, efficient HBB turns most of the $^{12}$C into $^{14}$N.
TDU is more efficient for compact, massive cores. HBB requires a more massive AGB star to reach sufficient temperatures at the base of the convective envelope to initiate the CNO cycle. Both processes are more efficient at low metallicity due to the lower opacities, hotter temperatures, and more compact internal structures. The details of each process are ultimately set by uncertain choices in stellar evolution models.

@@

\begin{table*}
	\centering
    \caption[]{For each AGB yield set, the IMF-averaged AGB C yield at solar metallicity $y_{\rm C, 0}^{\rm AGB}$, the fraction of solar C produced in the model $f_\odot^{\rm AGB}$, and the masses and metallicities each yield table is sampled on.
    $f_{\sun}^{\rm AGB}$ is calculated based on $y_{\rm C, \sun}^{\rm AGB}$ assuming the fiducial total C yield of $\Yct = 26.7\times 10^{-4}$.
    }

	\label{tab:agb}
    \begin{tabular}{c  c  c p{4cm} p{4cm}} % four columns, alignment for each
		\hline 
        AGB table 
                & $y_{\rm C, \sun}^{\rm AGB}\times10^4$ %& y
                &  $f_{\sun}^{\rm AGB}$
                & masses ($\Mo$) & metallicities ($Z$)\\
        \hline
        \fruity 
                & 3.11
                & 0.12
                & 1.3, 1.5, 2, 2.5, 3, 4, 5, 6
                & 0.0001, 0.0003, 0.001, 0.002, 0.003, 0.006, 0.008, 0.01, 0.014, 0.02
                \\
        \aton 
                & -0.009
                & -0.0003
                & 1.5, 2, 2.5, 3, 3.5, 4, 4.5, 5, 6, 6.5, 7
                & 0.0003, 0.001, 0.002, 0.004, 0.008, 0.014, 0.04
                \\
        \monash 
                &  2.96
                & 0.11& 1, 1.25, 1.5, 1.75, 2.25, 2.5, 2.75, 3, 3.25, 3.5, 3.75, 4, 4.5, 5, 5.5, 6, 7 
                & 0.0028, 0.007, 0.014, 0.03
                \\
        \nugrid 
                & 9.25
                &  0.35
                & 1, 1.65, 2, 3, 4, 5, 6, 7
                &  0.0001, 0.001, 0.006, 0.01, 0.02
                \\
		\hline
	\end{tabular}
\end{table*}

@@

CCSN models predict a wide range of C yields, spanning nearly a factor of ten, even at fixed metallicity.
Fig.~\ref{fig:y_cc} shows the integrated yields of massive star models from the literature. 
 The \cite{LC18} models, which include rotation, show that the induced mixing (e.g., \citealt{frischknecht+16}) can dramatically increase the magnitude and metallicity dependence of $\Ycc$. As we will later empirically show, CCSN C production needs to be strongly metallicity-dependent at $Z \approx \Zo$, which is consistent with \citepos{LC18} rapidly rotating models and to a lesser extent \citet{NKT13}.
Fig.~\ref{fig:y_cc} also shows the IMF-averaged AGB star yields from the \fruity\ model (in Fig.~\ref{fig:agb-ssp}). Especially at $Z\approx \Zo$, most CCSN models dominate AGB C production.

@@

The metallicity dependence of CCSN C yields is due to an interplay between rotation, opacity, and stellar evolution. As opacity increases with metallicity, massive stars experience stronger winds. In particular, C-enriched envelope material is lost through winds before conversion into heavier elements, resulting in metallicity-dependent C production \citep{LC18}.
While there are many stellar models providing predictions of CCSN yields, the results of these models are highly uncertain. Furthermore, rotation, binarity, and explodability introduce substantial uncertainties in CCSN predictions \citep{farmer+21}.

@@

SN Ia produce negligible C. Of the models from \citet{iwamoto+99}, \citet{seitenzahl+13}, and \citet{gronow+21a, gronow+21b}, the maximum C production is less than 4 per cent of our required solar C yield, and most likely less than 2 per cent (assuming  $2.2\times 10^{-3}$ SN Ia explosions per solar mass of star formation; \citealt{maoz+mannucci12}).
Novae are important contributors of $^{13}$C, but do not process enough material to impact $^{12}$C production \citep{starrfield+20}. $^{13}$C is beyond the scope of this work and accounts for a small fraction of the C in the Universe.

@@

Fig.~\ref{fig:c_evo} shows evolutionary tracks in \caah{} and \caafe{} in the fiducial model at 1\,kpc intervals of radius.
The \caah\ trend reaches an equilibrium state within \about{5}\,Gyr. \caafe\ tracks continue to evolve due to the long tail of the SN Ia DTD.

The evolution of chemical abundances can be understood as the interplay between different processes with different time-scales. 
 Initially, CCSN dominate enrichment. The [C/Mg] abundance ratio is set by the relative yield $\Ycc/\Ymg$. Due to the metallicity dependence of $\Ycc$, [C/Mg] increases with  [Mg/H]. [Mg/H], [Fe/H], and [C/H] all rise rapidly. [Mg/Fe] is initially high due to the delayed production time-scales of SN Ia Fe, as expected from the analytic models of \citet{WAF17}.
 Of order 100 Myr later, delayed sources enrich the ISM. AGB stars release C, raising [C/Mg], and SN Ia expel Fe, lowering [Mg/Fe]. 
 Several Gyr later, [Mg/H] reaches equilibrium. As the metal abundance stabilizes, so do metallicity-dependent CCSN yields, allowing C to reach equilibrium. [Fe/H] continues to rise slowly due to ongoing SN Ia from old stellar populations. 
 Finally, the system reaches approximate equilibrium and evolution slows. \

@@

\caption[]{The median low-$\alpha$ [C/N] ratio as a function of [Mg/H] for our fiducial model (blue line), compared with APOGEE subgiants (medians as black points with 16th and 84th percentile ranges as grey bars). Stars are binned into 20 equal-number bins. Combining our results with the suggested N yield from \citet{james+23} \ (rescaled to our adopted yields) explains the thin-disk evolution of both C and N. 
}
\label{fig:nitrogen}
\end{figure}

@@

    \caption[]{
        Trends in [C/Mg] with [Mg/H] (left) or [Mg/Fe] (right). 
        Stars are binned into 20 (left) or 12 (right) equal-number bins. 
        The left panel only shows low-$\alpha$ stars, and the right panel shows stars where $-0.15\leq {\rm [Mg/H]} \leq -0.05$.
        Coloured lines represent the median [C/Mg] in bins of [Mg/H] or [Mg/Fe] for each model.
        Black points and grey dashes represent the median and 16th-84th percentiles of [C/Mg] in each bin in the \citet{jack}~sample. 
        The total solar C yield is held fixed in each model. 
        %
        \textbf{Top}: Models with different slopes of the metallicity dependence of the CCSN yield, $\zcc$ (Eq.~\ref{eq:y_cc}), where higher $\zcc$ corresponds to a more rapid increase in C production with metallicity. \textbf{Middle}: Models with different fractions of solar C produced in AGB stars, $\fagb$. 
        \textbf{Bottom:} Models where both $\fagb$ and $\zcc$ are adjusted simultaneously. 
    }
    \label{fig:zeta_f}
\end{figure*}

@@

\section{Model parameters and measurement scatter} \label{sec:extra_scatter}

\begin{table}
	\centering
        \caption[]{Description of the yields used for models presented in this paper (see Sections \ref{sec:agb}, \ref{sec:vice}, and \ref{sec:results} for details).}
	\label{tab:model_parameters}
